\documentclass[bsc]{abdnthesis}

%% For citations, I would recommend natbib for its                          
%% flexibility, particularly when named citation styles are used, but                
%% it also has useful features for plain and those of that ilk.                      
%% The natbib package gives you the following definitons                             
%% that extend the simple \cite:                                                     
%   \citet{key} ==>>                Jones et al. (1990)                              
%   \citet*{key} ==>>               Jones, Baker, and Smith (1990)                   
%   \citep{key} ==>>                (Jones et al., 1990)                             
%   \citep*{key} ==>>               (Jones, Baker, and Smith, 1990)                  
%   \citep[chap. 2]{key} ==>>       (Jones et al., 1990, chap. 2)                    
%   \citep[e.g.][]{key} ==>>        (e.g. Jones et al., 1990)                        
%   \citep[e.g.][p. 32]{key} ==>>   (e.g. Jones et al., p. 32)                       
%   \citeauthor{key} ==>>           Jones et al.                                     
%   \citeauthor*{key} ==>>          Jones, Baker, and Smith                          
%   \citeyear{key} ==>>             1990                                             

\usepackage[round,colon,authoryear]{natbib}
\usepackage[acronym]{glossaries}
\setlength{\bibsep}{0pt}
\bibliographystyle{apalike}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage{listings}

\usepackage[T1]{fontenc}

\title{Towards Evaluating Creativity in Language}
\author{Matey Krastev}
% IMO this is a bit silly, but some like to include these. To remove,
% delete this declaration and remove the option from the
% \documentclass definition above.
%\qualifications{PhD, Computer Science, University College London, 1997\\%            
%BEng (Hons.) Electrical and Electronic Engineering, The University of Wales, Swansea, 1992}
\school{Department of Computing Science}

%%%% In the final submission of a thesis, this should only be the year
%%%% of submission.  However, it is useful to use \date{\today} for drafts so that
%%%% they don't get mixed up.
    
\date{2023}

%% It is useful to split the document up as chapters and include
%% them.  LaTeX will sort out all the numbering and cross-referencing
%% for you --- if you run it enough times!

%% If you want to include only a couple of chapters then use the
%% \includeonly{} command with a list of the file/chapter names that
%% you wish to include.  NB, this must be in the preamble.

% \includeonly{introduction,faq}

% \def\sfthing#1#2{\def#1{\mbox{{\small\normalfont\sffamily #2}}}}

% \sfthing{\PP}{P}
% \sfthing{\FF}{F}

%% This will make sure that all cross-references are correct (including
%% references to those file not included) but will produce a dvi
%% file with only those files/chapters you specify included.

\input{sections/abbreviations.tex}

% Commands for making in-line notes
\usepackage{xcolor}
% \newcommand{\ajs}[1]{\textcolor{orange}{[AJS: #1]}}
\newcommand{\ajs}[1]{}
% \newcommand{\note}[1]{\textcolor{red}{[#1]}}
\newcommand{\note}[1]{}
% \newcommand{\mk}[1]{\textcolor{blue}{[MK: #1]}}
\newcommand{\mk}[1]{}

\begin{document}
 
%%%% Create the title page and standard declaration.

\maketitle
\makedeclaration

%%%% Then the abstract and acknowledgements
\begin{abstract}

    We hypothesize that what we call creativity equates to an expression of a given author's unique character through their work, be it in text, speech, art, and so on. Furthermore, we claim that those can be represented as a certain set of linguistic features that can be found in text, e.g. more of this feature makes this text seem more creative.

    To that end, we explore the field of linguistic creativity through the lens of statistical text analysis. We investigate a variety of statistical measures that have been suspected to aid in text classification tasks, genre identification and others. Furthermore, we explore prior work in psycholinguistics aiming to measure the effect of words on the reader, and how those effects can be used by authors to the effect of solidifying an author's style. We also hypothesize a set of novel measures that can be linked with the understanding of structure in text. 
    
    In this work, we implement a system for working with text data, as well as methods for applying the measures we have explored. Then, we carry out an experimental evaluation of the system on a set of texts using the implemented metrics, on a broad variety of applicable NLP tasks, such as genre classification (where we accomplish 99.9\% accuracy across 3 categories of text), authorship identification (achieved 55.4\% accuracy for 50 authors, and 22\% accuracy for 1000 authors), and machine-generated text detection (above 70\% accuracy). 
    
    We hypothesize that the measures we have explored can be used to evaluate the creativity of text -- creativity in terms of a distinguishing feature for an author -- and how this lays out some groundwork for further research into the deep topic of computational creativity.
\end{abstract}

\begin{acknowledgements}
    Many thanks to my academic supervisor, Dr. Arabella Sinclair, for constantly being involved with the project, guiding me during my first steps in Natural Language Processing, bearing through my ramblings and entertaining my ideas.

    Many thanks to everyone who asked me questions about the project, which served to clarify and potentially solidify my ideas, as well as for giving me the opportunity to better explain my ideas to myself through explaining them to others. I will remember that one time I tried to explain this idea in terms of shepherding, sheep and shepherd dogs. 
    
    Many thanks go to the many educators who have taught me throughout the years, teaching me to think critically, rationally, and to put everything under scrutiny.
\end{acknowledgements}


%%%% It should have a table of contents, but delete the other two as
%%%% necessary.

\tableofcontents

\clearpage
\printnoidxglossary[type=\acronymtype,sort=standard, nonumberlist, title=Abbreviations]
\clearpage

%\listoftables
%\listoffigures

\include{sections/introduction}
% \include{faq}
\include{sections/background.tex}
\include{sections/methods.tex}
\include{sections/design.tex}
\include{sections/implementation.tex}
\include{sections/evaluation.tex}
\include{sections/discussion.tex}
% \include{sections/conclusion}

\appendix
% \include{proof}

\include{sections/user_manual.tex}
\include{sections/maintenance_manual.tex}

\bibliography{bibliography}

\end{document}
